### 2016-09-16 draft/draft0.tex
updates

%%% On-going present
%%
%% Beginning of file 'sample.tex'
%%
%% Modified 2005 December 5
%%
%% This is a sample manuscript marked up using the
%% AASTeX v5.x LaTeX 2e macros.

%% The first piece of markup in an AASTeX v5.x document
%% is the \documentclass command. LaTeX will ignore
%% any data that comes before this command.

%% The command below calls the preprint style
%% which will produce a one-column, single-spaced document.
%% Examples of commands for other substyles follow. Use
%% whichever is most appropriate for your purposes.
%%
%%\documentclass[12pt,preprint]{aastex}

%% manuscript produces a one-column, double-spaced document:

%%\documentclass[manuscript]{aastex}

%% preprint2 produces a double-column, single-spaced document:

 %\documentclass[preprint2]{aastex}
%\documentclass[preprint2]{aastex}
%\documentclass[manuscript]{aastex}
\documentclass[iop]{emulateapj}
\renewcommand{\baselinestretch}{0.95}
%% Sometimes a paper's abstract is too long to fit on the
%% title page in preprint2 mode. When that is the case,
%% use the longabstract style option.

%% \documentclass[preprint2,longabstract]{aastex}

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
%% If you are submitting to a journal that translates manuscripts
%% into SGML, you need to follow certain guidelines when preparing
%% your macros. See the AASTeX v5.x Author Guide
%% for information.

\usepackage{enumitem}

\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$~}\fi}} 
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

%% You can insert a short comment on the title page using the command below.

%\slugcomment{Not to appear in Nonlearned J., 45.}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.
%% Running heads will not print in the manuscript style.

%\title[]
%{}

\shorttitle{Reshift Dependent Volume effect}
\shortauthors{X.-D. Li, C. Park, C.G. Sabiu, H. Park, D.H. Weinberg, J. Kim, S.E. Hong}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Cosmological constraints from the redshift dependence of %geometric deformation
the galaxy angular 2-point correlation function}

%\author{Xiao-Dong~Li\altaffilmark{1}, Changbom~Park\altaffilmark{1}, J.~E.~Forero-Romero\altaffilmark{2} and Juhan Kim%\altaffilmark{3}}
%\affil{\altaffilmark{1}School of Physics, Korea Institute for Advanced Study, Heogiro 85, Seoul 130-722, Korea }
%\affil{\altaffilmark{2}Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, %Colombia}
%\affil{\altaffilmark{3}Center for Advanced Computation, Korea Institute for Advanced Study, 85 Hoegi-ro, Dongdaemun-%gu, Seoul 130-722, Korea }
%\affil{\it (xiaodongli@kias.re.kr, cbp@kias.re.kr, je.forero@uniandes.edu.co, kjhan@kias.re.kr)}

%\author[Xiao-Dong~Li, Changbom~Park, Cristiano G. Sabiu, Juhan Kim and Sungwook E. Hong]
%{ Xiao-Dong Li$^{1,\dagger}$, Changbom Park$^{1}$, Cristiano G. Sabiu$^{2}$, Juhan Kim$^{3,1,\star}$, Sungwook E. Hong$^{1}$\\
%$^1$School of Physics, Korea Institute for Advanced Study, 85 Heogi-ro, Dongdaemun-gu, Seoul 130-722, Korea\\
%$^2$Korea Astronomy and Space Science Institute, 776, Daedeokdae-ro, Yuseong-gu, Daejeon, 305-348, Korea\\
%$^3$Center for Advanced Computation, Korea Institute for Advanced Study, 85 Hoegi-ro, Dongdaemun-gu, Seoul 130-722, Korea\\
%$^{\dagger}$xiaodongli@kias.re.kr\\
%$\star$Corresponding Author: ***@kasi.re.kr}

\author{Xiao-Dong Li, }
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\author{Cheng Cheng, }
\affil{Kavli Institute for Theoretical Physics China, Institute of Theoretical Physics, Chinese Academy of Sciences, Zhong Guan Cun Street 55\#, Beijing, 100190, P.R. China}
\affil{University of Chinese Academy of Sciences, P.R. China}
\author{Changbom Park,}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\author{Hyunbae Park, Cristiano G. Sabiu, }
\affil{Korea Astronomy and Space Science Institute, Daejeon 305-348, Korea}
%\author{David H. Weinberg,}
%\affil{Department of Astronomy and CCAPP, The Ohio State University, 140 West 18th Avenue, Columbus, OH 43210, USA}
%\author{Donald P. Schneider,}
%\affil{Department of Astronomy and Astrophysics, The Pennsylvania State University, 
%University Park, PA 16802 }
%\affil{Institute for Gravitation and the Cosmos, The Pennsylvania State University, 
%University Park, PA 16802 }
\author{Juhan Kim\altaffilmark{1},}
\affil{Center for Advanced Computation, Korea Institute for Advanced Study, 85 Hoegi-ro, Dongdaemun-gu, Seoul 130-722, Korea}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\and
\author{Stephen Appleby, Sungwook E. Hong}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}

%\altaffiltext{1}{xiaodongli@kias.re.kr}
%\altaffiltext{2}{cbp@kias.re.kr}
%\altaffiltext{3}{csabiu@kasi.re.kr}
%\altaffiltext{4}{\bf Hyunbae: your email}
\altaffiltext{1}{Corresponding Author: kjhan@kias.re.kr}

% \email{xiaodongli@kias.re.kr}
% \email{cbp@kias.re.kr}
% \email{je.forero@uniandes.edu.co}

% \author{S. Djorgovski\altaffilmark{1,2,3} and Ivan R. King\altaffilmark{1}}
% \affil{Astronomy Department, University of California,
%     Berkeley, CA 94720}
%
% \author{C. D. Biemesderfer\altaffilmark{4,5}}
% \affil{National Optical Astronomy Observatories, Tucson, AZ 85719}
% \email{aastex-help@aas.org}
%
% \and
%
% \author{R. J. Hanisch\altaffilmark{5}}
% \affil{Space Telescope Science Institute, Baltimore, MD 21218}
%
% %% Notice that each of these authors has alternate affiliations, which
% %% are identified by the \altaffilmark after each name.  Specify alternate
% %% affiliation information with \altaffiltext, with one command per each
% %% affiliation.
%
% \altaffiltext{1}{Visiting Astronomer, Cerro Tololo Inter-American Observatory.
% CTIO is operated by AURA, Inc.\ under contract to the National Science
% Foundation.}
% \altaffiltext{2}{Society of Fellows, Harvard University.}
% \altaffiltext{3}{present address: Center for Astrophysics,
%     60 Garden Street, Cambridge, MA 02138}
% \altaffiltext{4}{Visiting Programmer, Space Telescope Science Institute}
% \altaffiltext{5}{Patron, Alonso's Bar and Grill}

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.

\begin{abstract}
We propose to use the redshift of the shape of galaxy angular 2-point correlation function (2pCF) to constrain cosmology.
The method is a continue of our series of works \citep{Li2014,Li2015,Li2016} probing the redshift dependence of Alcock-Paczynski effect and volume effect. 
A wrongly adopted cosmology result in redshift-dependent mis-scaling in the constructed galaxy distribution,
which leads to redshift-dependent rescaling in the measured angular 2pCF.
We test our method on the HR4 mock galaxies having 457, 406, 352, 206 and 228 million galaxies at redshfits 0, 0.5, 1, 1.5, 2,
and find the redshift evolution of the scaling of angular 2pCF is sensitive to the adopted cosmological parameters and 
relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the redshift space distortion.
Analyzing redshift evolution of the shape of 2pCF on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$ we derive tight constraints on $\Omega_m$ and $w$.
Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future photometric surveys of DES, LSST to derive tight cosmological constraints.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. The uncommented
%% example has been keyed in ApJ style. See the instructions to authors
%% for the journal to which you are submitting your paper to determine
%% what keyword punctuation is appropriate.

\keywords{large-scale structure of Universe --- dark energy --- cosmological parameters}

%% From the front matter, we move on to the body of the paper.
%% In the first two sections, notice the use of the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below. We have
%% chosen the first three characters of the first author's name plus
%% the last two numeral of the year of publication as our KEY for
%% each reference.

%% Authors who wish to have the most important objects in their paper
%% linked in the electronic edition to a data center may do so by tagging
%% their objects with \objectname{} or \object{}.  Each macro takes the
%% object name as its required argument. The optional, square-bracket
%% argument should be used in cases where the data center identification
%% differs from what is to be printed in the paper.  The text appearing
%% in curly braces is what will appear in print in the published paper.
%% If the object name is recognized by the data centers, it will be linked
%% in the electronic edition to the object data available at the data centers
%%
%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
%% the brackets must be escaped with backslashes when used in the first
%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
%%  Otherwise, LaTeX will issue an error.

\section{Introduction}

Dark energy has become the most important scientific issue since the discovery of cosmic acceleration \citep{Riess1998,Perl1999}.
Many theoretical and observational effect has been made \citep{Weinberg,Li2011,DHW2013}, but its properties still remain far from accurately known and its nature still remains mystery. 
%The simplest and currently favored solution is the additional field called dark energy which has negative pressure and is spatially homogeneous.
%The simplest explanation for dark energy is that it is a cosmological constant or vacuum energy, with constant equation of state (EoS) $w = -1$.

To measure the properties dark energy we need accurate probe of the cosmic expansion history. 
In large scale structure surveys the two representative geometrical observables are the angular diameter distance $D_A$ and the Hubble factor, $H$. 
If these can be precisely measured as functions of redshift, then tight constraints can be placed on the parameters of the cosmological model,
e.g., the matter density $\Omega_m$ and the equation of state (EoS) of dark energy, $w$.

Assuming an incorrect cosmological model for the coordinate transformation from redshift space to comoving space leaves residual geometric distortions. 
These distortion include volume effect which is the alteration of the comoving volume size of structures,
and the Alcock-Paczynski (AP) effect, induced by the fact that distances along and perpendicular to the line of sight are fundamentally different.
%Measuring the ratio of galaxy clustering in the radial and transverse directions, provides a probe of the Alcock-Paczynski (AP) effect \citep{AP1979}.
%There have been a series of works studying the application of the AP test to the LSS
A series of statistically method has been proposed to measure the AP test in the LSS
\citep{Ballinger1996,Matsubara1996,Marinoni2010,Ryden1995,LavausWandelt1995}
and have been applied to a series of LSS surveys \citep{Outram2004,Blake2011,ChuangWang2012,
Reid2012,Beutler2013,Linder2013,2014arXiv1407.2257S, Jeong2014,Sutter2014,2014ApJ...781...96L,
Alam2016, Beutler2016, Sanchez2016,Qingqing2016}. 
The probe of volume effect is usually probed through number counting of 
galaxy clusters \citep{PS1974,VL1996}, topology \citep{topology};
it is also part of the physical effects measured in the measurements of the BAO scale \citep{some reference},
or the probe of the shape of 2pCF and power spectrum \citep{some reference}.

In \cite{Li2014,Li2015,Li2016} we proposed a series of methods to probe the geometric information encoded in the LSS data.
We make use of the fact that the AP or volume effect commonly introduce redshift-dependent geometric distorion in the LSS,
which can be distinguished from the other effects such as the redshfit space distortion (RSD).
RSD arises because the radial distances of galaxies are inferred from redshifts which is contaminated by the pecular motion of galaxies, 
leading to apparent anisotropy even if the adopted cosmology is correct \citep{Ballinger1996}.
RSD is the usually main caveats for probe geometry using LSS,
and we found it can be largerly avoid by focusing on the the redshift dependence of the anisotropy.

\cite{Li2014} proposed used the {\it galaxy density gradient field} 
to characterize the anisotropies in LSS and tested the idea on Horizon Run 3 (HR3) N-body simulations \citep{horizonrun},
showing the redsift evolution of the gardient field can be used to infer unbiased constraints cosmological parameters and is insensitive to RSD.
The same topic was revisited in \cite{Li2015}, but using the statistical tool of {\it galaxy two-point correlation function} (2pCF) as a function of angle, $\xi(\mu)$.
When incorrect cosmological parameters are adopted, the shape of $\xi(\mu)$ appears anisotropic due to the AP effect, 
and the amplitude is shifted by the change in comoving volume;
both effects have significant redshift dependence.
The RSD effect, although significantly distorting $\xi(\mu)$, 
exhibits much less redshift evolution compared to the effect coming from incorrectly adopted cosmologies.
The method was applied to BOSS DR12 galaxies \citep{Li2016} to probe the redshift dependent of AP effect and lead to tight cosmological constraints.

We developed a method measuring the redshift of the shape of galaxy angular correlation function to constrain cosmology.
A wrongly adopted cosmology result in redshift-dependent mis-scaling in the constructed galaxy distribution,
which leads to redshift-dependent change in the angular 2pCF shape.
The signal is sensitive to the adopted cosmological parameters and relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the RSD effect.
We test our method on the HR4 mock galaxies having hundreds of million galaxies at redshfits 0-2.
Analyzing redshift evolution of the shape of 2pCF on scales $5 - 40 h^{-1} {\rm Mpc}$, we derive tight constraints on $\Omega_m$ and $w$.

In this paper we continue our series of work and probe to use the angular correlation function as a statistical tool to probing the volume effect in the angular direction.
%The method of redshift dependent AP effect has been shown great power to constrain cosmological parameters \cite{Li2014,Li2015,Li2015}.
%We hope this method also.
%The two methods probe different information encoded in the 2pCF and can be combined to fully probe the geometrical effect in LSS.
%Or one can also use the redshift evolution of 2D 2pCF xi(s,mu) to have all information included.
%We developed a method measuring the redshift of the shape of galaxy angular correlation function, $\hat \omega_{r_\perp}$, to constrain cosmology.
A wrongly adopted cosmology result in redshift-dependent mis-scaling in the constructed galaxy distribution,
which shifts apparent scale of the clustering patterns;
this manifests itself as a redshift-dependent rescaling of the measured angular correlation function.
We find this signal is sensitive to the adopted cosmological parameters and relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the RSD effect.
We test our method on the HR4 mock galaxies.
The method can be applied to current and future LSS surveys,
especially, the photometric surveys such as DES and LSST which probe a large amount of galaxies and the large redshift error will not quite affect the angular 2pCF measurement.

The outline of this paper proceeds as follows. 
In \S 2 we briefly review the nature and consequences of the geometric distortion in LSS when performing coordinate transforms in a cosmological context. 
In \S 3 we describe the mock galaxies produced from N-body simulations which are used to test our methodology.
The methodology is presented in \S 4.
We conclude in \S 6.

\section{Volume Effect in a Nutshell}
\label{sec:Voleffect}

\begin{figure*}
   \centering{
   \includegraphics[height=8cm]{fig1_a.eps}
   \includegraphics[height=8.5cm]{fig1_b.eps}
   }
   \caption{\label{fig_xyquan}
   %Apparent distortion of objects in four wrongly assumed cosmologies, assuming a true cosmology of .
   The scaling in four wrongly assumed cosmologies ..., % $\Omega_m=0.41$, $w=-1.3$ and $\Omega_m=0.11$, $w=-0.7$, 
   assuming a true cosmology of $\Omega_m=0.26$, $w=-1$.
   Left panel shows a series objects distributes as five perfect squares, measured by an observer located at the origin.
   Their true positions and shapes are plotted in blue dashed lines.
   The observer measures the redshifts of these objects, adopts the wrong cosmologies to compute the distance,
   and obtains wrong results (the red solid lines).
   %The red solid lines show their positions and shapes when the wrong cosmologies 
   % are assumed to compute distance from measured redshift.
   %The mis-estimation of angular diameter distance lead to wrong 
   Right panel shows the mis-estimation of angular diameter distance when the wrong cosmologies are adopted.
   %In our mock surveys we split the samples at $z=0.3$, 0.6, 0.9 and 1.2, as marked by the vertical lines.
   %measured by an observer located at the origin.
   %For reference, blue dashed squares show their shapes and positions in the true cosmology.
   }
   % Need a better plot to show angular size variation!!!
\end{figure*}

%Considering we are Let us consider an object in the universe with certain size and shape.
%Its observed redshift span $\Delta z$ and angular size $\Delta \theta$ are related with the comoving sizes by
In this section we briefly introduce the scaling effect caused by wrongly assumed cosmological parameters.
A more detailed description has been provided in \cite{Li2014,Li2015,Li2016}.

Suppose that we are probing the size of some objects in the Universe.
We measure its redshift span $\Delta z$ and angular size $\Delta \theta$,
then compute its sizes in the radial and transverse directions using the following formulas
\begin{equation}\label{eq:distance}
\Delta r_{\parallel} = \frac{c}{H(z)}\Delta z,\ \ \Delta r_{\perp}=(1+z)D_A(z)\Delta \theta,
\end{equation}
where $H$ is the Hubble parameter and $D_A$ is the angular diameter distance.
In the particular case of a flat universe composed by a cold dark matter component and a constant EoS dark energy component, they take the forms of
\begin{eqnarray}\label{eq:HDA}
& &H(z) = H_0\sqrt{\Omega_ma^{-3}+(1-\Omega_m)a^{-3(1+w)}},\nonumber\\
& &D_A(z) = \frac{1}{1+z}r(z)=\frac{1}{1+z}\int_0^z \frac{dz^\prime}{H(z^\prime)},
\end{eqnarray}
where $a=1/(1+z)$ is the cosmic scale factor,
$H_0$ is the present value of Hubble parameter and $r(z)$ is the comoving distance.

In case wrong values of $\Omega_m$ and $w$ are adopted, 
the inferred $\Delta r_{\parallel}$ and $\Delta r_{\perp}$ are wrong,
resulting in wrong estimation in the object's shape (AP effect) and size (volume effect).
These effects and their cosmological consequences have been investigated in \cite{Li2014,Li2015,Li2016}.

In this paper we focus on the mis-estimation of the angular size $\Delta r_{\perp}$. %scaling of angular size,
The ratio of mis-estimation is
\begin{equation}\label{eq:DA}
 %{\rm Degree\ of\ volume\ magnification \equiv}
 %{\rm rat}_{\rm volume\ mag}\equiv
 \alpha_{\perp} \equiv \frac{\Delta r_{\perp, \rm wrong}}{\Delta r_{\perp, \rm true}}
 = \frac{D_{A, \rm wrong}}{D_{A, \rm true}},
\end{equation}
where ``true'' and ``wrong'' denote the values of $D_A$ in the true cosmology and wrongly assumed cosmologies, respectively.

An illustration is provided in the left panels of Figure \ref{fig_xyquan}.
Suppose that the true cosmology is a flat $\Lambda$CDM with present matter ratio $\Omega_m=0.26$
and standard dark energy EoS $w=-1$.
If we were to distribute a series of perfect squares at various distances ranging from 500 Mpc/h to 3\,000 Mpc/h,
and an observer located at the origin were to measure their redshifts and infer the sizes of the squares
using the distance-redshift relations of four incorrect cosmologies
\begin{eqnarray}
 & &{\rm (i)}.\ \Omega_m=0.40,\ w=-1.0, \nonumber \\ 
 & &{\rm (ii)}.\ \Omega_m=0.15,\ w=-1.0, \nonumber\\ \noindent
 & &{\rm (iii)}.\ \Omega_m=0.26,\ w=-0.5,\nonumber\\ \noindent
 & &{\rm (iv)}.\ \Omega_m=0.26,\ w=-1.5,  \nonumber \noindent 
\end{eqnarray}
as a result, 
the shapes of the squares appear distorted (AP effect),
and their sizes are wrongly estimated (volume effect).
Cosmological models (i,iii) yield to compressed size (in both angular and LOS direction),
and the degree of compression increases with increasing distance;

The mis-estimation of angular size, ${\Delta r_{\perp, \rm wrong}}/{\Delta r_{\perp, \rm true}}$, 
are displayed in the right panel of Figure \ref{fig_xyquan}.
All cosmologies, ${\Delta r_{\perp, \rm wrong}}/{\Delta r_{\perp, \rm true}}$ evolves a lot in the redshift range $0<z<2$.
As an example, when adopting the quintessence cosmology $\Omega_m=0.26,\ w=-0.5$,
the angular size is underestimated by 
%12.5\%, 17.7\%, 20.1\%, 21.3\%
8.9\%, 12.3\%, 13.6\%, 14.1\% 
at $z=0.5,1,1.5,2$.

%Thus, when wrong cosmological parameters are adopted,
%the apparent size of objects are distorted, and the magnitude of distortion also varies with redshift.

In sum, as a consequence of incorrectly adopted cosmologies, 
the size of the objects is mis-estimated and the magnitude of mis-estimation depends on the redshift.
In this paper we use the galaxy angular 2pCF to probe the mis-estimation of angular size ${\Delta r_{\perp, \rm wrong}}/{\Delta r_{\perp, \rm true}}$.

\section{The Simulation Data}\label{sec:data}

We test the method using the mock galaxy samples produced by the Horizon Run 4  (HR4) N-body simulation \citep{hr4,hong2016}.

HR4 was made within a cube of volume $({3.15}\ h^{-1}\rm Gpc)^3$ using  $6300^3$ particles with mass $m_p\simeq 9 \times 10^9 \hMsun$.
The simulation adopted the second order Lagrangian perturbation theory (2LPT) initial conditions at $z_{i}=100$
and a WMAP5 cosmology $(\Omega_{b},\Omega_{m},\Omega_\Lambda,h,\sigma_8,n_s)$  = (0.044, 0.26, 0.74, 0.72, 0.79, 0.96) \citep{komatsu 2011}.
%The particle mass is $\simeq m_{p} \simeq 9.02 \times 10^9 \hMsun$.
%This starting redshift, combined  with 2LPT initial conditions, ensures an accurate mass function and power spectrum \citep{2014NewA...30...79L}.

Mock galaxies are produced from the simulation based on a modified one-to-one correspondence scheme \citep{hong2016}. 
The most bound member particles (MBPs) of simulated halos are adopted as tracers of galaxies.
The merger timescale is computed to get the lifetime of merged halos.
Merger trees of halos are constructed by tracking their MBPs from $z = 12$ to 0;
when a merger event occurs, the merger timescale is computed using the formular of \cite{jiang2008} to 
determine when the satellite galaxy is completely disrupted.
%We modeled the position and the velocity of a galaxy from the position and velocity of its corresponding MBP.
%By using the abundance matching, 
%we modeled the luminosity of a central/isolated galaxies from their current mass
%and of satellite galaxies from their mass at the time of infall.

The resulting mock galaxies was found to reproduce the 2pCF of SDSS DR7 volume-limited galaxy sample \citep{zehavi2011} very well \citep{hong2016}.
The mock galaxies shows a similar finger of god (FOG) feature \citep{FOG} as the observation.
The projected 2pCF of the mock and observational samples agree within 1$\sigma$ CL
on scales greater than 1 ${h^{-1}}$Mpc.

%As will be introduced in the next section, 
%The output of HR4 simulation includes one all-sky light cone mock galaxy catalogue reaching $r=3\,150$ ${h^{-1}}$Mpc 
%and a series ({\bf how many?}) of snapshot mock galaxy catalogues at different redshifts.
In this analysis we use snapshot data of HR4 galaxies at $z=0,0.5,1,1.5,2$.
%We require subhalos to have at least 30 member particles, 
%so the minimal mass of galaxies is $\simeq m_{p} \simeq 2.7 \times 10^{11} \hMsun$.
Setting a minimal halo mass of $3\times 10^{11} \hMsun$, 
%$4.57\times 10^8$, $4.06\times 10^8$, $3.52\times 10^8$, $3.06\times 10^8$ and $2.28 \times 10^8$ mock galaxies,
we select 457, 406, 352, 206 and 228 million mock galaxies at the five redshifts,
corresponding to a number density of 
%$1.46\times 10^{-2}$, $1.30\times 10^{-2}$, $1.13\times 10^{-2}$, $0.98\times 10^{-2}$ and $0.73\times 10^{-2}\ (h^{-1} \rm Mpc)^{-3}$,
1.46, 1.30, 1.13, 0.98 and 0.73 in unit of $ 10^{-2} h^{3} \rm Mpc^{-3}$,
%$1.46\times 10^{-2}$, $1.30\times 10^{-2}$, $1.13\times 10^{-2}$, $0.98\times 10^{-2}$ and $0.73\times 10^{-2}\ (h^{-1} \rm Mpc)^{-3}$,
respectively.
By applying a uniform mass cut, at higher redshift we are selecting less galaxies with larger galaxy bias.

As an illustration, Figure \ref{fig_scatter} displays the 2D distribution of a subsample of mock galaxies in five redshifts,
with $x$, $y$ coordinates computed using the ``correct'' values of cosmological parameters $\Omega_m=0.26,\ w=-1$ (left panels) 
and a wrong set of cosmological parameters $\Omega_m=0.05,\ w=-1.5$ (right panels), respectively.

When the correct cosmology is adopted the cosmic scale is correctly estimated in five redshifts.
The only factor leads to evolution of galaxy distribution with redshift is the gravitational growth of structure.
With the decreasing the redshift, we clearly see how the clusters and filaments form and grow.
On the other hand, when the wrong cosmology is adopted, there is an artifical scaling of scales in the constructed map.
The separations among galaxies are over-estimated by 
$25.6\%,47.3\%,62.2\%,71.7\%$ at redshifts of $0.5,1.0,1.5,2.5$,
leading to a clear evolution of the sizes of structures with redshift.

The growth of structure leads to growth of clustering strength of structures on all scales,
while the volume effect maintain the clustering pattern and uniformly re-scale all the structures.
Their imprints on the large scale structure is different and should be distinguishable in the statistical analysis.
In the following, we show that they affect the angular correlation function measurements in different ways, 
and can be easily separated.

%The growth of structure also make the size of structures evolves with redshift.
%In the next subsection we will discuss how to distinguish these two effects in the galaxy angular 2pCF.

%The left panels of Figure \ref{fig_scatter} show the $x,y$ coordinates of galaxies in a 
%$260 h^{-1} {\rm Mpc} \times 130 h^{-1} {\rm Mpc} \times 105 h^{-1} {\rm Mpc}$ volume, 
%taken from the five snapshots.
%The gravitational growth of structures with decreasing redshift is clearly shown.
%We display all mock galaxies with $M> 3.0 \times 10^{11} \hMsun$.
%At lower redshift, galaxies become more massive, so they are more galaxies satisfying the threshold.

%If one adopts a wrong cosmology to compute galaxy positions, there would be a scaling of the galaxy distribution.
%In the left panel we show the case when a wrong set of parameters $\Omega_m=0.05,\ w=-1.5$ is assumed,
%leading to large upscaling of $25.6\%,47.3\%,62.2\%,71.7\%$ at redshifts $z=0.5,1.0,1.5,2.5$.

%Compared with the true galaxy distribution, the apparent distribution inferred by wrong set of parameters
%show clear evolution with redshift due to the different scaling at different redshifts.
%In case of Figure \ref{fig_scatter}, the apparent size of large scale structures increases with redshift.

The growth of structure also make the size of structures evolves with redshift.
In the next subsection we will discuss how to distinguish these two effects in the galaxy angular 2pCF.

\begin{figure*}
   \centering{
   %\includegraphics[height=5.9cm]{fig8_0.eps}
   %\includegraphics[height=5.9cm]{fig8_wrongcos_0.eps}
   %\includegraphics[width=16cm]{fig2.eps}
   %\includegraphics[width=16cm,natwidth=8,natheight=14]{fig2.pdf}
   \includegraphics[width=16cm]{fig2.eps}
   }
   \caption{\label{fig_scatter}
  blah
   }
   % Need a better plot to show angular size variation!!!
\end{figure*}

\section{Methodology}

%\subsection{Galaxy distribution at different redshifts}

We use the angular 2pCF as a statistical tool to probe the volume effect.
The galaxy 2pCF as a function of galaxy separation in the angular direction, $\omega(r_\perp)$, is computed for snapshot data of mock galaxies at five redshifts.
We adopt the Landy-Szalay estimator~\citep{1993ApJ...412...64L},
\begin{equation}
\omega(r_\perp)=\frac{DD-2DR+RR}{RR}\ ,
\end{equation}
where $DD$ is the number of galaxy--galaxy pairs, 
$DR$ the number of galaxy-random pairs, 
and $RR$ is the number of random--random pairs, 
all separated by a distance defined by $r_\perp\pm\Delta_{r_\perp}$ where we choose $\Delta_{r_\perp}=1h^{-1}\rm Mpc$.
The random catalogue consists of unclustered points uniformly distributed in a same size box. 
In an effort to reduce the statistical variance of the estimator, we use 50 times as many random points as we have galaxies.

Considering the large number of galaxies and random points the 2pCF is computed part by part in subsamples with 
size of 1575 $h^{-1}{\rm Mpc}\times 1575\ h^{-1}{\rm Mpc}\times 105\ h^{-1}{\rm Mpc}$.
The Z direction with thickness $105\ h^{-1}\rm Mpc$ is treated as the radial direction (the $r$ direction) 
and the X-Y directions with each of width $1575 h^{-1}\ \rm Mpc$ are the angular plane.
For our simulation box size we can have 120 such subsamples.
The average of the measurements in all subsamples is adopted as the 2pCF of the whole sample,
while the covariances of them are adopted as the covariance matrix after multiplying a factor of $1/\sqrt{119}$.

\subsection{Galaxy angular 2pCF: amplitude and shape at different redshifts}\label{sec_2pCF_diffz}

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig3.eps}
    \includegraphics[width=18cm]{fig3_b.eps}
   }
   \caption{\label{fig_diffz}
  2pCF at different redshfits.
   }
\end{figure*}

The upper-left panel of Figure \ref{fig_diffz} displays the angular 2pCF measured from HR4 mock galaxies. % (multiplied by scale $r$) in the five snapshot.
We multiply $\omega$ by the separation $r_\perp$ to have similar amount of statistical uncertainty on all scales.

At different redshifts, we see large evolution in the amplitude of $\omega(r_\perp)$.
The amplitude of 2pCF is proportional to the clustering strengh which is affected by the gravitational growth of structures and the bias of the galaxies.
The amplitude is highest at $z=0$ where the large scale structures experienced most gravitational growth,
and also gets higher at $z>1$ with increasing redshift as a reason of more biased galaxies at higher redshift.
%leading to stronger clustering strength and higher amplitude of 2pCF.
%On the contrary, at higher redshift we find $\omega(z=2)>\omega(z=1.5)>\omega(z=1)$,
%The trend is different from $z=1,1.5,2$,
%simply because by applying a uniform mass cut we are selecting more biased galaxies at higher redshift.

Different from the large variation of amplitude among the different redshifts, %The shape of $r_\perp\omega$ also evolves with redshift.
the shape of $r_\perp\omega$ maintains similar at all redshifts;
it always peaks at $r\sim9 h^{-1}\rm Mpc$ and monotonically drops at larger or smaller scales.
The only exception is the small enhancement at $r\lesssim 2 h^{-1}\rm Mpc$,
which is caused by the non-linear growth of structures on small scales 
and is much more significant at lower redshfit.

In order to directly compare the shape of 2pCF at different redshifts,
in the middle panel of Figure \ref{fig_diffz} we show the $r_\perp\omega$ normalized by the overall amplitude within $5h^{-1}{\rm Mpc} < r < 50h^{-1}{\rm Mpc}$
(here after $\hat \omega_{r_\perp}$):
\begin{equation}
 \hat \omega_{ r_\perp} \equiv  \frac{ {r_\perp\omega}(r_{\perp})}{ \int_{r_{\perp, \rm min}}^{r_{\perp, \rm max}}r_{\perp} \omega(r_{\perp}) d r_{\perp} / (r_{\perp,\rm max}-r_{\perp,\rm min}) },
\end{equation}
where we choose $r_{\perp, \rm min}=5 h^{-1}$ Mpc, $r_{\perp, \rm max}=40 h^{-1}$ Mpc in this analysis.
Scales of $r_{\perp, \rm min}<5 h^{-1}$ Mpc are affected by non-linear growth of strucutre, 
the mis-estimation of whose contribution may lead to systematic errors in the derived cosmological constraints.
On scales larger than $40 h^{-1}$ Mpc it is may be possible to have precise analytically modelling the shape of the 2pCF
 \citep{Salvador2014,Salvador2016} which should be able to extract more cosmological information than our method 
(although out method should is also applicable).

The nice overlapping of the five redshift results
justify the small evolution of shape with redshift.
In the right panel, we further show the residual evolution of $\hat r_\perp\omega$ at high redshifts 
with respect to $z=0$.
There is only 1-4\% relative enhancement 
at $r<10h^{-1}\rm Mpc$,
and $<1.5\%$ relative suppression at $r>25 h^{-1}\rm Mpc$.
The trend is monotonic with redshift.

There is one point should be emphasized here.
In real observations we are observing {\it different} objects in different redshfit shells.
To mimick this we comparing slices of different redshifts at {\it different} locations.
As an exammple, if at $z=0$ we take $\hat\omega_{r_\perp}$ measured within $0h^{-1}{\rm Mpc}<Z<105 h^{-1}{\rm Mpc}$,
at higher redshifts we then adopt measurement within $105h^{-1}{\rm Mpc}<Z<210 h^{-1}{\rm Mpc}$ for a comparision.
%and take the difference between them to get $\delta \hat{r_\perp\omega}$.
%So we are always comparing 2pCF measured from different samples of galaxies.
If simply comparing the 2pCF of a {\it same} subsample of galaxies at different redshifts, 
one would significantly underestimate the statistical undertainty of $\delta \hat{\omega_{r_\perp}}$ by ignoring the cosmic variance.
The error bars and covariance are always estimated in this way to take the cosmic variance into consideration.
%\footnote{In all figures of this paper The error bars displayed in all figures are estimated in this way; 
%from the 120 subsamples and for the error bar of redshift evolution of 2pCF 
%we always take the cosmic variance into consideration.}.

As shown by Figure \ref{fig_scatter}, the gravitational growth of strucutre enhance the clustering strengh of structures on all scales;
here it manifests itself as an enhanced amplitude of 2pCF at low redshift.
The shape of 2pCF, which represents the {\it relative strength of clustering among different scales},
maintain similar with redshift except the non-linear scales of $r_\perp < 5 h^{-1}\rm Mpc$.
%mainly increases the amplitude of

%In all, we conclude that the shape of the 2pCF is more robust than the amplitude against redshift and galaxy bias.

\subsection{Galaxy angular 2pCF: Cosmological Effect }

The measured angular correlation function of galaxies with positions constructed using the wrong cosmology $\Omega_m=0.05,\ w=-1.5$
is displayed in the lower panels of Figure \ref{fig_diffz}.
The mis-scaling uniformly shifts clustering patterns on all scales, lead to a wrong estimation of $\hat \omega_{r_\perp}$ 
related with the correct one as
\begin{equation}
 \hat \omega_{r_\perp,\rm wrong}(r) = \hat \omega_{r_\perp, \rm correct}(\alpha_{\perp} r),
\end{equation}
a simple consequence of the phenomenon that clustering patterns at distance $r$ now appears on $\alpha_{\perp} r$.
The redshift evolution of $\alpha_{\perp}$ leads to redshfit evolution of shape.

%Lower panels of Figure \ref{fig_diffz} shows the $r_\perp\omega$ in five 
%snapshots, in case that the cosmology $\Omega_m=0.05,\ w=-1.5$ 
%is adopted to construct the galaxy distribution (the right panels of Figure \ref{fig_scatter}).
The change in shape as well as the redshift evolution is displayed in the middle panel.
The upscaling of comoving distance lead to a stretched shape.
With the increasing of redshift,
the peak location is shifted from $9h^{-1} \rm Mpc$ to $15 h^{-1}\rm Mpc$ at $z=2$,
as a result of the fact that the angular separation is upscaled by $71.7\%$.
Correspondingly, the right panel shows that, 
compared with $\hat \omega_{r_\perp}$ at $z=0$ there is a 20-40\% change at higher redshifts.

The mis-scaling not only changes the of the 2pCF but also modifies its amplitude. %The amplitude of $r_\perp\omega$ is amplified due to the stretch of scale.
as shown in the left panel, due to the stretch of scale the amplitude is enhanced at higher redshifts;
the higher the redshift, the more the enhancement.%the amplitude monotonically increases as a function of redshift.

Although the change of amplitude could be a more significant consequence of wrong cosmology than the alteration of shape,
it is mixed with other factors; 
the growth of structure and a larger galaxy bias can also lead to a stronger clustering and thus an enhanced amplitude of 2pCF.
%it is more difficult to 
%But this phenomenon can also appear in case of gravitational growth of structure or increasing of bias,
%so we will not make use of it to do cosmological constraint.
In the analysis to extract cosmological information, we just make use of the redshift evolution of the 2pCF shape, 
which is less affected by factors of structure growth and galaxy bias.
%which is as a signal suggesting that the assumed cosmological parameters are wrong.

\subsection{Systematic effects}

Sec. \ref{sec_2pCF_diffz} shows that $\hat \omega_{r_\perp}$ measured from a set of constant mass cut samples at different redshifts is close to each other.
The gravitational growth of structure will not have a large impact on $\hat \omega_{r_\perp}$ on scales $\gtrsim5 h^{-1}$ Mpc.
The other factors may introduce redshift evolution of the clustering of measured galaxies include the 
redshift space distortion (RSD), galaxy bias, redshift error, and other properties of galaxies, such as the mass, morphology, color, concentration. 
Here we test two ``major'' systematical effects, the galaxy bias and the RSD effect.

%The same as \cite{Li2014,Li2015,Li2016} we need to correct for the redshift evolution of $r_\perp\omega$ caused by effects other than the cosmological effect.

%Figure \ref{fig_diffz} already shows that, the growth of structure, especially in low redshift and on non-linear scales, 
%changes the shape of $r_\perp\omega$ and caused relative enhancement on $r\lesssim2 h^{-1}\rm Mpc$.
%So we limit the region of scale to $r>5h^{-1}\rm Mpc$ to minimize its effect.

\subsubsection{Galaxy Bias}

Galaxies are biased tracers of dark matter field;
more massive galaxies reside in regions with higher density contrast.
In large scale structure surveys, the bias of the galaxy sample is an important fator affecting on the clustering properties of the sample.

We check the effect of galaxy bias on $\hat \omega_{r_\perp}$.
The upper panels of Figure \ref{fig_sys} displays the 2pCF measured from galaxies within a $3175\times3175\times105$ $(h^{-1}\rm Mpc)^3$ volume, taken from the $z=0$ snapshot.
Four values of minimal mass limits, $3\times 10^{11} M_{\odot}$, $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$,
are imposed to create three samples with different galaxy bias.
The measured angular 2pCF are displayed.

The different mass cuts result in large variation in the amplitude of the 2pCF (left panel),
while the shape of the 2pCF remains less affected (middle panel).
Compared with the sample of $M>3\times 10^{11} M_{\odot}$,
samples with mass limits of $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$
have amplitude of 2pCFs enhanced by $10\%,\ 50\%$ and $100\%$,
while the change in $\hat\omega_{r_\perp}$ is only $0.5\%,\ 2\%$ and $4\%$.
So the effect of galaxy bias is much avoided by utilizing the shape of the 2pCF rather than the amplitude.

\subsubsection{Redshift Space Distortion}

The galaxy peculiar velocity contributes to the observed redshift and distorted the inferred galaxy radial position,
known as the redshift space distortion.
It is the major systematics in our previous works \citep{Li2014,Li2015,Li2016}
where we derive cosmological constraints from the redshift evolution of the 3D galaxy distribution.

The effect of RSD is expected to be much milder in this work. 
The angular positions of the galaxies is not shifted by RSD at all;
the only effect of RSD come in the procedure of sample preparation. %preparing the subsamples of galaxies for 2pCF analysis.
The galaxies observed in a survey are split into shells of subsamples with different redshift ranges for the purpose 2pCF measurement.
RSD disortes the galaxy redshift and as a result some galaxies (especially those close to the boundaries of shells) could be classified to wrong redshift shells.

The lower panel of Figure \ref{fig_sys} displays the RSD effect on 2pCF.
To have the effect of RSD we shift the radial coordinates of galaxies according to the relation 
\begin{equation}\label{eq:zvpeu}
\Delta z = (1+z) \frac{v_{{\rm Z}}}{c},
\end{equation}
where $v_{{\rm LOS}}$ is the galaxy peculiar velocity in the radial direction.
This lead to mis-classfication of some galaxies when we split the box into slices.
%Then we split the whole box into 30 slices with thickness $105h^{-1}\rm Mpc$, 
%and measure the angular 2pCF.

Comparing the plot with Figure \ref{fig_diffz} one clearly see the effect of RSD.
The amplitude of measured 2pCF is enhanced by $\sim 10\%$ in case of considering the RSD effect.
The slop of $\hat\omega_{r_\perp}$ slightly suppressed.
While the redshift evolution of $\hat\omega_{r_\perp}$ is still small...
Although the pattern is a little modified...
So RSD is not a big issue here... {\bf after anna jobs finished we have better plots and re-revise this section }
%If looking at the redshift evolution,
%the curves $\hat{r_\perp\omega} - \hat{r_\perp\omega}(z=0)$ displayed in the right panel,
%they changed a lot compared with the case of no RSD effect (the right panel of Figure \ref{fig_diffz}),
%but do not become much larger. %But RSD effect does not introduce large redshfit evolution, although the shape of  is

Similar to RSD, redshift errors of galaxies (typically 0.02 for photometric survey) also results in fuzzy boundaries of shells and mis-classfication of galaxies.
This effect should also be properly quantified in real observations.

%In \cite{Li2014,Li2015,Li2016}, the systematic effects contributing to the redshfit evolution of galaxy clusterings 
%are modeled in mock surveys and subtracted.
%Similar treatment could be conducted here.
%One shall construct mock surveys with the above effects included for the correction of systematics.

%The outer boundary of the mock survey becomes fuzzy due to the peculiar velocity effect 
%on the galaxy redshifts (Eq. (\ref{eq:zvpeu})).
%A population of galaxies, that are expected to enter the $z<0.7$ region from the outside, is missing.
%(these $z>0.7$ galaxies have a peculiar velocity toward us and 
%their observed redshfits become lower than 0.7).
%To avoid this problem we set the maximal redshift at 0.693, 
%23.3 $h^{-1}{\rm Mpc}$ away from $z=0.7$.

%3.000e11', '1.000e12', '4.000e12',  '1.000e13

\begin{figure*}
   \centering{
   \includegraphics[width=16cm]{fig4_a.eps}
   \includegraphics[width=16cm]{fig4_b.eps}
   }
   \caption{\label{fig_sys}
  Systematic effects.
   }
\end{figure*}

\subsection{Likelihood Analysis}

We design the likelihood function to quantify
the redshift evolution of $\hat\omega_{r_\perp}$.
A cosmology resulting in smaller evolution is considered to be more likely the correct one.
The procedure here is very similar to what we did in \cite{Li2014,Li2015,Li2016}.
We comapre the high redshift $\hat\omega_{r_perp}$ and the lowest redshift measurement:
\begin{equation}\label{eq:chisq1}
\chi^2\equiv \sum_{i=2}^{n_{z}} \sum_{j_1=1}^{n_{r}} \sum_{j_2=1}^{n_{r}} {\bf p}(z_i,r_{j_1}) ({\bf Cov}_{i}^{-1})_{j_1,j_2}  {\bf p}(z_i,r_{j_2}),
\end{equation}
where $n_z$ is the number of redshifts which is 5 in this analysis, 
$n_r$ is number of binning in $\hat{\omega_{r_\perp}}(r_\perp)$,
which is 35 since we have $r_\perp$ bins 
with width $1h^{-1}\rm Mpc$ in a range of $5h^{-1}{\rm Mpc}\leq r\leq 40h^{-1}{\rm Mpc}$.
${\bf p}(z_i,r_{j})$ is the redshift evolution of the angular correlation function shape with systematic effects subtracted
\begin{eqnarray}\label{eq:bfp}
 {\bf p}(z_i,r_{j}) \equiv&\ \delta \hat{r_\perp\omega}(z_i,z_1,r_j) - \delta \hat{r_\perp\omega}_{\rm sys}(z_i,z_1,r_j)
\end{eqnarray}
${\bf Cov}_i$ is the covariance matrix estimated from the ${\bf p}(z_i,r_{j})$ measured from 120 subsamples.
%It is difficult to conduct the 2pCF of the whole snapshot box with hundreds of millions of galaxies.
%We split the box into 120 subsamples with $X,Y,Z$ sizes of $1575h^{-1}{\rm Mpc},1575h^{-1}{\rm Mpc}, 105h^{-1}{\rm Mpc}$
%and measure the angular 2pCF in the $X-Y$ plan in these subsamples. 
%The average of the 2pCFs are taken as the 2pCF of the whole box.
%The covariance matrix is also estimated from the redshift evolution of these subsamples.
As mentioned before, for a robust estimation of covariance matrix, 
we always compare slices at different locations to 
include the cosmic variance.

The covariance matrix inferred from a finite number of samples
is always a biased estimate of the true matrix \citep{Hartlap}.
This can be corrected by rescaling the inverse covariance matrix as 
\begin{equation}
 {\bf Cov}^{-1}_{ij,\rm Hartlap} = \frac{N_s - n_{r}-2}{N_s-1} {\bf Cov}^{-1}_{ij},
\end{equation}
where $N_s=120$ is the number of mocks used in covariance estimation.
For the options of this analysis the rescaling is as large as 1.43.
In the case that one have 2\,000 mocks the rescaling is less than 1.02.

%estimated from the 72 sets of PSB mock galaxies identified from HR3 N-body simulation.

Figure \ref{fig_cosmo} displays the 2pCF and the likelihood values in case of adopting 
the four wrong cosmologies of Figure \ref{fig_xyquan}.
For the cosmologies $\Omega_m=0.4,w=-1$ and $\Omega_m=0.26,w=-0.5$,
the compression of structure shift the large scale clustering patterns into small scales,
and also becomes more significant at higher redshift.
We get 2pCFs with steeper slope and the higher redshift the steeper.
For cosmologies $\Omega_m=0.15,w=-1$ and $\Omega_m=0.26,w=-1.5$,
the stretch of structure leads to shallower slope of $\hat{\omega_{r_\perp}}$ and the higher the redshift the shallower.
In all cosmologies we have significant detection of redshift evolution.
We computed the $\chi^2$ values according to Equation \ref{eq:chisq1}, 
and found these cosmological parameters are strongly disfacored 
at $\gtrsim30\sigma$ CL.

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig5.eps}
   }
   \caption{\label{fig_cosmo}
   2pCF in cosmologies.
   }
\end{figure*}

\section{Cosmological constraint}

We constrain $\Omega_m$ and $w$ through Bayesian analysis (\cite{Bayesian}; also see \cite{LB2002,Li2016}).
We assue the likelihood takes the form
\begin{equation}
 \mathcal{L} \propto \exp\left[-\frac{\chi^2}{2}\right]
\end{equation}
and scan the parameter space in $\Omega_m-w$ plane to obtain the 68.3\% and 97.4\% CL regions.
The result is displayed in Figure \ref{fig_contours}.

\begin{figure*}
   \centering{
   %\includegraphics[width=7cm]{Tpcf--contour.eps}
   %\includegraphics[width=16cm]{Tpcf--contour-others.eps}
   %\includegraphics[width=16cm,natwidth=52,natheight=40]{fig10.eps}
   \includegraphics[width=16cm]{fig9.eps}
   }
   \caption{\label{fig_contours}
   Likelihood contours (68.3\%, 95.4\%) in the $\Omega_m-w$ plane from our method.
   }
\end{figure*}

We get tight constraint on the two paramters.
The 2$\sigma$ contour lies within the region $0.23<\Omega_m<0.285$, $-1.1<w<-0.9$.
The thin shape of contour means, when combining with the another observational data with different direction of degeneracy (e.g. CMB),
very tight combined constraint can be obtained.
In case of fixing one parameter at its best-fit value and infer the statistical uncertainty of the other one,
one will obtain 1$\sigma$ uncertainty of $\delta\Omega_m\approx0.002,\delta w\approx0.01$.

\section{Concluding Remarks}

We developed a method measuring the redshift of the shape of galaxy angular correlation function, $\hat \omega_{r_\perp}$, to constrain cosmology.
A wrongly adopted cosmology result in redshift-dependent mis-scaling in the constructed galaxy distribution,
which leads to redshift-dependent rescaling of $\hat \omega_{r_\perp}(r_\perp)$.
The signal is sensitive to the adopted cosmological parameters and relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the RSD effect.
We test our method on the HR4 mock galaxies having 457, 406, 352, 206 and 228 million galaxies at redshfits 0, 0.5, 1, 1.5, 2;
analyzing redshift evolution of the shape of 2pCF on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$ we derive tight constraints on $\Omega_m$ and $w$.

The angular correlation function is adopted as a statistical tool to probing the scaling effect in the angular direction.
One can also use the 2D 2pCF as a function of both angular and radial scale to fully probe the scaling effect in 3D galaxy distribution.

There are already works using the shape of angular 2pCF to constrain cosmology
by directly comparing the theoretical 2pCF to the observed one \citep{Salvador2014,Salvador2016}.
While our method just utilize the fact the the shape of angular 2pCF is insensitive to redshift in correct cosmology.
Our method is complementary to these works in the sense that,
we do not require accurate analytic modeling and should be applicable on smaller scales.
The two different approaches can be combined together to fully extract the cosmological information.

The method of redshift dependent AP effect has been shown great power to constrain cosmological parameters \cite{Li2014,Li2015,Li2015}.
We hope this method also.
The two methods probe different information encoded in the 2pCF and can be combined to fully probe the geometrical effect in LSS.
Or one can also use the redshift evolution of 2D 2pCF xi(s,mu) to have all information included.

The method is designed to be applied to current and future LSS surveys,
with a particular emphasis to the photometric surveys which probe a large amount of galaxies and the large redshift error will not quite affect the angular 2pCF measurement.
An accurate modeling of redshfit error, galaxy bias and RSD effects could be crucial for the successful application of the method.
DES will probe 300 million galaxies at $z<1.4$, and LSST will probe 20 billion galaxies at $z<2$.
We hope our method play an active role in  deriving cosmological constraints from LSS surveys.

\section*{Acknowledgments}

We thank the Korea Institute for Advanced Study for providing computing resources (KIAS Center for Advanced Computation Linux Cluster System).
%We thank Seokcheon Lee and Graziano Rossi for many helpful discussions.
We would like to thank Yi Zheng for useful discussions.
This work was partially supported by the
Supercomputing Center/Korea Institute of Science and
Technology Information with supercomputing resources
including technical support (KSC-2013-G2-003).

\appendix

\

\begin{thebibliography}{}

\bibitem[Ade et al. (2015)]{Planck2015}
Ade, P.A.R., Aghanim, N., \& Arnaud, M., et al. arXiv:1502.01589

\bibitem[Alam et al.(2016)]{Alam2016}
Alam, S., Ata, M., \& Bailey, S., et al. 2016,
submitted to MNRAS (arXiv:1607.03155)

\bibitem[{{Alam} {et~al}\mbox{.}(2015{\natexlab{a}}){Alam}, {Albareti},
  {Allende Prieto}, {Anders}, {Anderson}, {Anderton}, {Andrews}, {Armengaud},
  {Aubourg}, {Bailey}, \& et~al.}]{dr12}
{Alam} S., Albareti, F.D.,\& Allende Prieto, C., {et~al.}, 2015,  ApJS, 219, 12

\bibitem[Alcock \& Paczynski(1979)]{AP1979}
Alcock, C., \& Paczynski, B. 1979, Nature, 281, 358

%\bibitem[Anderson et al.(2012)]{2012MNRAS.427.3435A} 
%Anderson, L., Aubourg, E., Bailey, S., et al.\ 2012, MNRAS, 427, 3435

\bibitem[Anderson et al.(2013)]{Anderson2013}
Anderson, L., Aubourg, \'E., \& Bailey, S. et al. 2014, MNRAS, 441, 24

%\bibitem[Bassett et al.(2002)]{Bassett2002}
%Bassett, B.A., Kunz, M., Silk, J., \& Ungarelli, C. 2002, MNRAS, 336, 1217

\bibitem[Ballinger, Peacock \& Heavens 1996]{Ballinger1996}
Ballinger, W.E., Peacock, J.A., \& Heavens, A.F. 1996, MNRAS, 282, 877

\bibitem[Betoule et al.(2014)]{JLA}
Betoule, M., Kessler, R., \& Guy, J., et al. 2014, A\&A, 568, 32

\bibitem[Beutler et al.(2011)]{6dFGS}
Beutler, F., Blake, C., \& Colless, M., et al. 2011, MNRAS, 416, 3017

\bibitem[Beutler et al.(2013)]{Beutler2013}
Beutler, F., Saito, S., \& Seo, H.-J., et al. 2013, MNRAS, 443, 1065

\bibitem[Beutler et al.(2016)]{Beutler2016}
Beutler, F., Seo, H.-J., \& Saito, S., et al. 2016,
arXiv:1607.03150

\bibitem[Blake et al.(2011)]{Blake2011}
Blake, C., Glazebrook, K., \& Davis, T. M., 2011, MNRAS, 418, 1725

\bibitem[Blake et al.(2013)]{WiggleZtopoloy}
Blake, C., James, J.B., \& Poole, G.B. 2013, MNRAS, 437, 2488

\bibitem[Bolton et al.(2012)]{Bolton2012}
Bolton, A.S., Schlegel, \& D.J., Aubourg E., et al. 2012, AJ, 144, 144

\bibitem[Boylan-Kolchin et al.(2008)]{B08}
Boylan-Kolchin, M., Ma, C.-P., \& Quataert, E. 2008, MNRAS, 383, 93

%\bibitem[Bueno Belloso et al. (2012)]{BB2012}
%Bueno Belloso, A., Pettinari, G.W., Meures, N., \& Percival, W.J. 2012, Phys. Rev. D, 86, 023530

%\bibitem[Chevallier \& Polarski(2001)]{CP2001}
%Chevallier, M., Polarski, D. 2001, Int. J. Mod. Phys. D, 10, 213

%\bibitem[Choi et al.(2010)]{choi 2010}
%Choi, Y.-Y., Park, C., Kim, J., Gott, J.R., 
%Weinberg, D.H., Vogeley, M.S., \& Kim, S.S. 2010, ApJS, 190, 181

\bibitem[Christensen et al.(2001)]{Bayesian}
Christensen, N., Meyer, R., Knox, L., \& Luey, B. 2001, Class. Quant. Grav., 18, 2677

%\bibitem[Chuang et al.(2013)]{Chuang2013}
%Chuang, C.-H., Prada, F., Beutler, F., et al. 2013, arXiv:1312.4889

\bibitem[Chuang \& Wang(2012)]{ChuangWang2012}
Chuang, C.-H., \& Wang, Y. 2012, MNRAS, 426, 226

%\bibitem[Corasaniti \& Copeland(2003)]{Corasaniti2003}
%Corasaniti, P.S., Copeland, E.J. 2003, Phys. Rev. D, 67, 063521

%eBOSS: 
%http://arxiv.org/abs/1508.04473
\bibitem[Dawson et al.(2015)]{eBOSS}
Dawson, K.S., Kneib, J.P., \& Percival, W.J., et al. 2015, accepted AJ

\bibitem[Dawson et al.(2012)]{Dawson et al. 2012}
Dawson, K.S., Schlegel, D.J., \& Ahn, C.P., et al. 2012, AJ, 145, 10

\bibitem[Efstathiou (2014)]{E14H0}
Efstathiou, G. 2014, MNRAS, 440, 1138

\bibitem[Eisenstein et al.(2011)]{Eisenstein et al. 2011}
Eisenstein, D.J.,  Weinberg, D.H., \& Agolet, E., et al. 2011, AJ, 142, 72

\bibitem[Feldman, Kaiser \& Peacock (1994)]{1994ApJ...426...23F} 
Feldman, H.A., Kaiser, N., \& Peacock, J.A.\ 1994, ApJ, 426, 23

\bibitem[Fukugita et al. (1996)]{Fukugita1996}
Fukugita, M., Ichikawa, T., \& Gunn, J.E., et al. 1996, AJ, 111, 1748
%Publication:	
%Astronomical Journal v.111, p.1748

%\bibitem[Gingold \& Monaghan(1977)]{GM1977}
%Gingold, R.A., \& Monaghan, J.J. 1977, MNRAS, 181, 375

%\bibitem[Gott et al.(2009)]{gott 2009}
%Gott, J.R., Choi, Y.-Y., Park, C., \& Kim, J. 2009, ApJ, 695, L45

%\bibitem[Gott et al.(2008)]{gott 2008}
%Gott, J.R., Hambrick, D.C., Vogeley, M.S., Kim, J., Park, C., Choi, Y.-Y.,
%Cen, R., Ostriker, J.P., \& Nagamine, K. 2008, ApJ, 675, 16

\bibitem[Gunn et al. (1998)]{Gunn1998}	
Gunn, J.E., Carr, M., \& Rockosi, C. et al. 1998, AJ, 116, 3040

\bibitem[Gunn et al.(2006)]{Gunn et al. 2006}
Gunn, J.E., Siegmund, W.A., \& Mannery, E.J., et al. 2006, AJ, 131, 2332

\bibitem[Guzzo et al.(2008)]{Guzzo2008}
Guzzo, L., Pierleoni, M., \& Meneux, B., et al. 2008, Nature, 451, 541

\bibitem[Hartlap et al.(2006)]{Hartlap}
Hartlap J., Simon P. \& Schneider P. [astro-ph/0608064].

\bibitem[Hong et al.(2016)]{hong2016}
Hong, S.E., Park, C.,\&  Kim, J. 2016, ApJ, 823, 103

\bibitem[Jackson (1972)]{FOG}
Jackson, J., 1972, MNRAS, 156, 1

\bibitem[Jennings et al.(2011)]{Jennings2011}
Jennings, E., Baugh, C.M., \& Pascoli, S. 2011, MNRAS, 420, 1079

%\bibitem[Jeong et al.(2014)]{Jeong2014}
%Jeong, D., Dai, L., Kamionkowski, M., \& Szalay, A.S. 2014, arXiv:1408.4648

\bibitem[Jiang et al.(2008)]{jiang2008}
Jiang, C.Y., Jing, Y. P., \& Faltenbacher, A., et al. 2008, ApJ, 675, 1095

\bibitem[Kaiser (1987)]{Kaiser1987}
Kaiser, N. 1987, MNRAS, 227, 1

\bibitem[Kim \& Park(2006)]{kim and park 2006}
Kim, J., \& Park, C. 2006, ApJ, 639, 600

\bibitem[Kim et al.(2009)]{2009ApJ...701.1547K} 
Kim, J., Park, C., Gott, J.R., III, \& Dubinski, J.\ 2009, ApJ, 701, 1547

\bibitem[Kim et al.(2015)]{hr4}
Kim, J., Park, C., L'Huillier, B., \& Hong, S. E. 2015, JKAS, 48, 213

\bibitem[Kim et al.(2011)]{horizonrun}
Kim, J., Park, C., Rossi, G., Lee, S.M., \& Gott, J.R. 2011, JKAS, 44, 217

\bibitem[Kitaura et al.(2015)]{MDPATCHY}
Kitaura, F.S., Rodrı\'{i}guez-Torres, S., Chuang, C.-H., et al. arXiv:1509.06400

\bibitem[Komatsu et al.(2011)]{komatsu 2011}
Komatsu, E., Smith, K. M., \& Dunkley, J., et al. 2011, ApJS, 192, 18

\bibitem[Lacey \& Cole(1993)]{LC93}
Lacey, C., \& Cole, S. 1993, MNRAS, 262, 627

\bibitem[Landy \& Szalay(1993)]{1993ApJ...412...64L} 
Landy, S.D., \& Szalay, A.S.\ 1993, ApJ, 412, 64

%EUCLID:
%http://arxiv.org/abs/1110.3193
\bibitem[Laureijs et al.(2011)]{EUCLID}
Laureijs, R., Amiaux, J., \& Arduini, S., et al. 2011, arXiv:1110.3193

\bibitem[Lavaux \& Wandelt(2012)]{LavausWandelt1995}
Lavaux, G., \& Wandelt, B.D. 2012, ApJ, 754, 109

%\bibitem[Levi et al.(2013)]{2013arXiv1308.0847L} 
%Levi, M., Bebek, C., Beers, T., et al.\ 2013, arXiv:1308.0847

\bibitem[Lewis \& Bridle (2002)]{LB2002}
Lewis, A., \& Bridle, S. 2002, Phys. Rev. D, 66, 103511

\bibitem[L'Huillier et al.(2014)]{2014NewA...30...79L} 
L'Huillier, B., Park, C., \& Kim, J.\ 2014, New Astronomy, 30, 79

\bibitem[Li et al.(2011)]{Li2011}
Li, M., Li, X.-D., Wang, S., \& Wang, Y. 2011, Commun. Theor. Phys., 56, 525

\bibitem[Li et al.(2014)]{Li2014}
Li, X.-D., Park, C., Forero-Romero, J., \& Kim, J. 2014, ApJ, 796, 137

\bibitem[Li et al.(2015)]{Li2015}
Li, X.-D., Park, C., Sabiu, C.G., \& Kim, J. 2015, MNRAS, 450, 807

\bibitem[Li et al.(2016)]{Li2016}
Li, X.-D., Park, C., Sabiu, C.G., \& Kim, J. 2016, submitted to ApJ

%\bibitem[Linder(2003)]{Linder2003}
%Linder, E.V. 2003, Phys. Rev. Lett., 90, 091301

\bibitem[Linder et al.(2014)]{Linder2013}
Linder, E.V., Minji, O., Okumura, T., Sabiu, C.G., \& Song, Y.-S. 2014, Phys. Rev. D, 89, 063525

\bibitem[L{\'o}pez-Corredoira(2014)]{2014ApJ...781...96L} 
L{\'o}pez-Corredoira, M.\ 2014, ApJ, 781, 96

\bibitem[Mao et al. (2016)]{Qingqing2016}
Mao, Q., Berlind, A.A., Scherrer, R.J., et al. 2016, submitted to ApJ

\bibitem[Marinoni \& Buzzi(2010)]{Marinoni2010}
Marinoni, C., \& Buzzi, A. 2010, Nature, 468, 539

\bibitem[Matsubara \& Suto(1996)]{Matsubara1996}
Matsubara T., \& Suto, Y. 1996, ApJ, 470, L1

\bibitem[McCavana et al.(2012)]{M12}
McCavana, T., Micic, M., Lewis, G. F., et al. 2012, MNRAS, 424, 361

\bibitem[Morandi \& Sun (2016)]{MS2016}
Morandi, A., \& Sun, M. arXiv:1601.03741

\bibitem[Outram et al.(2004)]{Outram2004}
Outram, P.J., Shanks, T., Boyle, B.J., Croom, S.M., Hoyle, F., Loaring, N.S., 
Miller, L., \& Smith, R.J. 2004, MNRAS, 348, 745

%\bibitem[Parejko et al.(2013)]{Parejko2013}
%Parejko, J. K., Sunayama, T., Padmanabhan, N., et al. 2013, MNRAS, 429, 98

\bibitem[Parejko et al.(2013)]{Parejko2013}
Parejko J.K., et al., 2013, MNRAS, 429, 98

\bibitem[Parihar et al. (2014)]{CMASSLSS2014}
Parihar, P., Vogeley, M.S., \& Gott, J.R., et al. 2014, ApJ, 796, 86

\bibitem[Park et al.(2005)]{park 2005}
Park, C., Kim, J., \& Gott, J.R. 2005, ApJ, 633, 1

\bibitem[Park \& Kim(2010)]{topology}
Park, C., \& Kim, Y.-R. 2010, ApJL, 715, L185

\bibitem[Park et al. (2012)]{Park2012}
Park, C., Choi, Y.-Y., Kim, J., Gott, J.R., Kim, S.S., \&
Kim, K.-S. 2012, ApJ, 759, 7

\bibitem[Park et al. (2015)]{Park2015}
Park, C., Song, H., Einasto, M., Lietzen, H., \&
Heinamaki, P. 2015, JKAS, 48, 75

\bibitem[Peebles \& Ratra(2003)]{PR2003}
Peebles, P.J.E., \& Ratra, B. 2003, Reviews of Modern Physics, 75, 559

\bibitem[Percival et al.(2014)]{Percival2014}
Percival, W.J., Ross, A.J., \& S\'{a}nchez, A.G., et al. 2014, MNRAS, 439, 2531

\bibitem[Perlmutter et al.(1999)]{Perl1999}
Perlmutter, S., Aldering, G., \& Goldhaber, G., et al. 1999, ApJ, 517, 565

\bibitem[Press \& Shechter(1974)]{PS1974}
Press, W.H., \& Schechter, P.L. 1974, ApJ, 187, 425

\bibitem[Reid et al.(2012)]{Reid2012}
Reid, B., Samushia, L., \& White, M., et al. 2012, MNRAS, 426, 2719

\bibitem[Reid et al.(2016)]{Reidetal:2016}
Reid, B., Ho, S., \& Padmanabhan, N., et al.  2016, MNRAS, 455, 1553

\bibitem[Riess et al.(1998)]{Riess1998}
Riess, A.G., Filippenko, A.V., \& Challis, P., et al. 1998, AJ, 116, 1009

\bibitem[Riess et al.(2011)]{Riess2011}
Riess, A.G., Macri, L., \& Casertano, S., et al. 2011, ApJ, 730, 119
%A 3\% Solution: Determination of the Hubble Constant with the Hubble Space Telescope and Wide Field Camera

\bibitem[Ross et al.(2012)]{2012MNRAS.424..564R} 
Ross, A.J., Percival, W.J., \& S{\'a}nchez, A.G. et al.\ 2012, MNRAS, 424, 564

\bibitem[Ross et al.(2015)]{MGS}
Ross, A.J., Samushia, L., \& Howlett, C., et al. 2015, MNRAS, 449, 835

\bibitem[Ryden(1995)]{Ryden1995}
Ryden, B.S. 1995, ApJ, 452, 25

%\bibitem[Samushia et al.(2014)]{Samushia2014}
%Samushia, L., Reid, B. A., White, M., et al. 2014, MNRAS, 439, 3504

%\bibitem[Sanchez et al.(2013)]{Sanchez2013}
%Sanchez, A. G., Kazin, E. A., Beutler, F., et al. 2013, MNRAS, 433, 1202

%\bibitem[Sutter et al.(2014)]{Sutter2014}
%Sutter, P.M., Pisani, A., Wandelt, B.D., \& Weinberg, D.H. 2014, MNRAS, 443, 2983

\bibitem[Salvador et al. (2014)]{Salvador2014}
Salvador, S.A., S\^{a}nchez, A.G., Padilla, N.D., \& Baugh, C.M. 2014, MNRAS, 443, 2612

\bibitem[Salvador et al. (2016)]{Salvador2016}
Salvador, S.A., S\^{a}nchez, A.G., \& Grieb, J.N., et al. 2016, submitted to MNRAS

\bibitem[Sanchez et al.(2016)]{Sanchez2016}
Sanchez, A. G., Scoccimarro, R., \& Crocce, M., et al.
arXiv:1607.03147

\bibitem[Schlafly et al.(2010)]{Schlafly2010}
Schlafly E.F., Finkbeiner D.P., Schlegel D.J., et al. 2010, ApJ, 725, 1175

\bibitem[Schlafly \& Finkbeiner(2011)]{SF2011}
Schlafly E.F., \& Finkbeiner D.P. 2011, ApJ, 737, 103

%DESI:
%http://arxiv.org/abs/1106.1706
\bibitem[Schlegel et al.(2011)]{DESI}
Schlegel, D., Abdalla, F., \& Abraham, T., et al. 2011, arXiv:1106.1706

\bibitem[Smee et al.(2013)]{Smee2013}
Smee, S.A., Gunn, J.E., \& Uomoto, A., et al. 2013, AJ, 146, 32

\bibitem[Song et al.(2014)]{2014arXiv1407.2257S} 
Song, Y.S., Sabiu, C.G., 
Okumura, T., Oh, M., \& Linder, E.V.\ 2014, JCAP, 12, 005

\bibitem[Speare et al. (2015)]{Speare2015}
Speare, R., Gott, J.R., Kim, J., \& Park, C.
2015, ApJ, 799, 176

%\bibitem[Tojeiro \& Percival(2011)]{Tojeiro2011}
%Tojeiro R., \& Percivial W.J. 2011, MNRAS, 417, 1114

%\bibitem[Tojeiro et al.(2012)]{Tojeiro2012}
%Tojeiro, R., Percival, W. J., Wake, D. A., et al. 2012, MNRAS, 424, 136

\bibitem[Viana \& Liddle(1996)]{VL1996}
Viana, P.T.P., \& Liddle, A.R. 1996, MNRAS, 281, 323

\bibitem[Villalobos et al.(2013)]{V13}
Villalobos, \'{A}., ́De Lucia, G., Weinmann, S.M., Borgani, S., \& Murante, G. 2013, MNRAS, 433, L49

\bibitem[Weinberg (1989)]{SW1989}
Weinberg, S. 1989, Reviews of Modern Physics, 61, 1

\bibitem[Weinberg et al. (2013)]{DHW2013}
Weinberg, D.H, Mortonson, M.J., Eisenstein, D.J., et al. 2013, Physics Reports, 530, 87

\bibitem[White (2011)]{White2011}
White M., et al. 2011, ApJ, 728, 126

\bibitem[York et al.(2000)]{York et al. 2000}
York, D.G., Adelman, J., \& Anderson, J.E., et al. 2000, AJ, 120, 1579

\bibitem[Zehavi et al.(2011)]{zehavi2011}
Zehavi, I., Zheng, Z., \& Weinberg, D.H., et al. 2011, ApJ, 736, 59

\end{thebibliography}

\end{document}